---
title: "Applying louvain algorithm on PPI network to identify clusters/ communities of proteins interacting          with each other"
author: "Menna Arafat"
date:  "2024-8-31"
output: html_document
---




## create output directory
```{r}
dir.create("output")
```

## load packages

```{r}
## BiocManager::install("estrogen")
 suppressPackageStartupMessages({ 
library(igraph)
library(tidyr)
library(readxl)
library(openxlsx)
library(org.Hs.eg.db)
library(scales)
library(ggplot2)
library(tibble)
library(dplyr)
library(plyr)
 })
```

## load data

```{r}
list.files()
net= read.delim("data/ppi_1st_shell_50_blue.tsv")[,1:2]
net=net[!grepl("ENS*|^LOC", net$X.node1) & !grepl("ENS*", net$node2),]
```

##  TO integrate ppi with TF networks
```{r}
##  tf_net= read.csv("jaspar_grey.csv")[,c(1,3)]
##  tf= tf_net$ID
##  names(tf_net)= names(net)
##  net= rbind(net, tf_net)
## --------------------------------
## build igraph object from edge list
```

## create igraph object
```{r}
graph =graph_from_edgelist(net %>%as.matrix(), directed = FALSE)
```

## --------------------------------
## Get hub proteins by applying centrality measures
##  Degree Centrality, edges cconnected to a node

```{r}
degree_centrality <- igraph::degree(graph)  %>% sort(.,decreasing = T) 
print(names(degree_centrality )[1:5])
DC= names(degree_centrality)[1:5] %>% as.vector()
```

##  Betweenness Centrality, The number of shortest paths that pass through a vertex

```{r}
betweenness_centrality <- betweenness(graph, normalized = TRUE)  %>% sort(.,decreasing = T) 
print(names(betweenness_centrality)[1:5])
BC= names(betweenness_centrality)[1:5] %>% as.vector()
```

##  Eigenvector Centrality, The importance of a node based on the importance of its neighbors.

```{r}
eigenvector_centrality <- eigen_centrality(graph)$vector %>% sort(.,decreasing = T) 
print(eigenvector_centrality) %>% head()
eigc= names(eigenvector_centrality)[1:5] %>% as.vector()
```

## closeness centrality

```{r}
closeness_centrality= closeness(graph) %>% sort(.,decreasing = T)
closeness_centrality %>% head()
cc= names(closeness_centrality)[1:5] %>% as.vector()
```

##  PageRank,Nodes are considered important if they are linked to by other important nodes, with a damping factor accounting for random jumps.

```{r}
pagerank <- page_rank(graph)$vector %>% sort(.,decreasing = T) 
pagerank %>% head()
pr= names(pagerank)[1:5] %>% as.vector()

all_drivers= c(DC, BC, eigc, cc, pr ) %>% unique()
all_drivers
```

## visualization
##  col <- colorRampPalette(c("#F6E8C3", "#1FA187FF")) 
##  col(10)
##  Create a color vector, defaulting to black

```{r}
node_colors <- rep("gray80", vcount(graph))
```

##  Set the color of specific nodes to red

```{r}
node_colors[V(graph)$name %in% all_drivers] <- "orange"
## node_colors[V(graph)$name %in% tf] <- "#1FA187FF"

node_colors
```

## function to change label position (stackoverflow)

```{r}
radian.rescale <- function(x, start=0, direction=1) {
  c.rotate <- function(x) (x + start) %% (2 * pi) * direction
  c.rotate(scales::rescale(x, c(0, 2 * pi), range(x)))
}
n= vcount(graph)
lab.locs <- radian.rescale(x=1:n, direction=-1, start=0)


png("output/igraph_ppi_network.png", width=20000, height=20000, res= 1000)
plot(graph, edge.arrow.size=.5, vertex.color= node_colors, vertex.size=5, 
     vertex.frame.color="gray", vertex.label.color="black", 
     vertex.label.cex=1, vertex.label.dist= 1.2, vertex.label.degree=lab.locs,
     layout=layout_in_circle, #' ## layout_with_fr
     main="PPI Networks for blue Module") 

dev.off()
```

## ------------------------------
## Cluster modularity (louvain)

```{r}
clusterlouvain <- cluster_louvain(graph)
clusterleiden = cluster_leiden(graph, objective_function = "modularity", n_iteration=10)
```

## see the distance between clustering of two algorithms whaere small distance indicate hat those algorithms gave more similar results

```{r}
compare(clusterlouvain, clusterleiden)
```

## group1 <- V(graph)$name[which(clusterlouvain$membership == 1)]
## export communities

```{r}
communities= communities(clusterlouvain)
communities
communities= ldply(communities, rbind) %>% t()
communities[is.na(communities)]= ""
communities= communities %>% as.data.frame()
names(communities)= paste0("cluster", 1:length(communities))
communities
## write.xlsx(communities, "communities_louvain_blue.xlsx", rowNames= F)
```

## ----------------------
## to draw communities splitted for visualization

```{r}
coGrph <- delete_edges(graph, E(graph)[igraph::crossing(clusterlouvain, graph)])
plot(coGrph)
```

## ?crossing
## Stretch = 20

```{r}
col= grDevices::adjustcolor(c("pink", "#B0E0E6", "#BF812D", "#46337EFF"), alpha=0.3) 
```

#Apply enrichment analysis for each protein cluster then annotate clustere using the mostnlikely term/ biological process
## read the output file of functional enrichment using clusterprofiler

```{r}
terms= read.xlsx("data/communities_louvain_blue.xlsx", sheet= "term")
```

## adjust layout

```{r}
layout <- layout_with_kk(coGrph)
```

##  Calculate centroids for each community

```{r}
centroids <- sapply(1:4, function(i) {
  colMeans(layout[clusterlouvain$membership == i, ])
}) %>% t()

png("output/ppi_communities_louvain.png", width=14000, height=11800, res=1000)
plot(graph, layout=layout_with_kk(coGrph) ,
     vertex.color= clusterlouvain$membership, vertex.size=5, 
     vertex.label.color="black", 
     vertex.label.cex=1, vertex.label.dist= 1.2,
     mark.groups = communities(clusterlouvain), #' ## list(group1,group2, group3, group4),
     mark.shape=1, 
     mark.col=col,
     mark.border= col,
     mark.expand = 30)
text(x= c(centroids[,1][1]*.39, centroids[,1][2]*.1, centroids[,1][3]*.2, centroids[,1][4]*.15),
     y= c(centroids[,2][1]*.37, centroids[,2][2]*.3, centroids[,2][3]*.35, centroids[,2][4]*.31),
    labels = terms$term, col = "black", cex = 1.1)

dev.off()
```

## -----------------------------------------------------------------------------------
## -----------------------------------------------------------------------------------
## -----------------------------------------------------------------------------------
## -----------------------------------------------------------------------------------
###### play with some igraph functions


```{r}
graph[]
get.adjacency(net, attr="weight", sparse=F)

## all vertices
V(graph)$name


## all edges
E(graph)

## get neighbours
n= neighborhood(graph, order=1, V(graph)$name %in% c( "APCS", "SERPINA3","A1BG" ))
names(unlist(n))    


## get edges between certain node and another node, and their count/length         
E(graph)[ V(graph)[name=="SERPINA3" ] %--% V(graph)[name=="A1BG"] ] %>% length()   
```

